Boundary Labeling for annotated documents

### 2017-09-10 LaTeX/Paper.tex
Paper: Expanded terminology section and re-wrote other sections that used newly defined words in different contexts. Expanded Section 3. Program: No changes.

\section{Introduction}
Annotating a document is usually solved by adding footnotes or figures in an appropriate position and adding a simple reference in the text, leaving the reader to find the referenced content by themselves. Sometimes however, if a more obvious connection between the text and the referenced content is required, the reference is visibly connected to the text by drawing a straight line or a more complex path between them.

@@

\section{Terminology and Fundamentals} %Evtl in Introduction eingliedern?
While Boundary Labeling (or an equivalent concept) can be applied to a space with different geometry or more dimensions, this paper will only concern itself with two-dimensonal, euclidean space.
To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)%Add more figures for graph-terms, etc. or change

@@

\begin{itemize}
 \item \textbf{Graph:} A Set of Vertices and Edges (see below), often used to represent routing problems. Can be further classified depending on its properties.
 \item \textbf{Edge:} A connection between two vertices, usually represented as a line. Can be directional, limiting its usage, or have a cost (``Weight'') associated with its use.
 \item \textbf{Vertex:} A featureless object that can be connected to any number other vertices via edges. Usually depicted as a point or small shape. Commonly also referred to as \textbf{``Node''}, as will be the the case in this paper.
 \item \textbf{Label:} Additional information, usually represented as a box containing the additional information. Will also be referred to as \textbf{``Annotation''} in this paper.
 \item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.
 \item \textbf{Leader:} The path connecting the site to the label - depending on its shape, it can be described further using several subgroups which can be freely combined to form new leader types:
  \begin{itemize}
   \item \textit{S-Leader:} This Leader connects site and label in a straight line.
   \item \textit{O-Leader:} This leader runs orthogonally to the border between text and label area.
   \item \textit{P-Leader:} This Leader runs parallel to the border between text and label area. It must be combined with other leader styles, as it won't reach the label area otherwise.
  \end{itemize}
    For example, the leader from Fig.~\ref{fig:term} would be classified as an opo-Leader.
 \item \textbf{Port:} The location where the leader connects to the label. It may be restricted pre-determined locations, like only at the corners.
  \item \textbf{Monotonous/Monotonicity:} Describes a steady progression towards a direction or goal. Will be used in this paper to indicate that there's no loss of progress made from the site to the label, effectively restricting the direction a leader can go at each bend.
\end{itemize}
%Mehr Graphenterminologie? (zb: Path) +Polylinien Vertex/Node für Knoten in Polylinien bzw. Graphen -->Eindeutig!

@@

%Algorithmus(Ansatz) und Implementierung(noch immer abstrakt) trennen! Probleme, Details, ... gegen Ende!
\subsection{Algorithm}%''Underlying Algorithm''?
To easily create paths that exclusively use the space between words and lines, we decided to use a graph similar to the one Loose \cite{Loose2015} used, placing nodes above and below each word's beginning and end, and connecting them with their horizontally and vertically adjacent neighbours.%Umformulieren?
For the sites, we inserted an additional node above the center of the word, which will serve as the leader's starting point. We also decided to give the graph's edges a capacity of 1, which means that each leader will always use edges which haven't already be taken by other leaders. While this causes some sites to become un-routable due to another leader passing right through their starting node, it makes visualizing the leaders much easier. Our first approach was going through the labels in the order they appear in the text and placing each as far up as possible, to maximize the remaining space other labels could use.

Since putting the labels right next to the text would result in the labels' positions being restricted by the text's line spacing,we decided to leave a buffer zone in between the two areas, which allows the leaders to adjust between the position forced by the line spacing and the label's final position. As this allowed us to place the labels more freely, we adjusted the label's positions after all labels were placed to reduce the distance that needed to be covered in the buffer area.

\subsection{Implementation}
The program was written in Java, using only JGraphT\cite{JGraphT} as additional library. Since we only want to create leaders that don't intersect with the text, the graph was created alongside the placement of the words on the canvas, as it was easy to extract measurements at this point. Due to some problems with Java's various methods of calculating linebreaks, this process was done completely by hand. Since the Graph's nodes represent fixed locations on the canvas, they each have coordinates associated with them, with the nodes representing the sites containing extra information, such as references to the corresponding annotation and to the leader connecting the two (if existent).
%TODO: Code genauer durchlesen und erweitern!

%Similar to Loose's Approach\cite{Loose2015}, we used a Graph to model the available routing space for the algorithm. As a result of this approach, it is important to know exactly where each word is placed. While Java offers a function that automatically places text in lines of given width, it does so in a way that doesn't allow the content of each line to be read as a String afterwards, which makes it unusable for our purposes. Therefore each word must be placed manually, and to since it isn't the main focus of the program, each series of characters between a space is considered one word that has to be fully placed within one line.

%Using the dimensions of each word, we create the graph by creating and connecting nodes in the space between lines at the start and end of each word, and at the end of the text area. We also create special nodes for each annotated word, they are located above the word in a central position. Each node also is located at set co-ordinates, which allows the routing algorithms to easily fulfill the monotonicity constraint.
%We left some space in between the text area and the label area %Name=?(Lass dir was einfallen)
%to allow the labels to be placed more freely. While the routing graph only reaches up to the border of the text area, the connection between the graph and the label is done with an OPO-Leader. As there is the possibility of intersecting with other leaders, this means some extra adjustments have to be done. For demonstration purposes, we also included another algorithm that uses S-Leaders for this section.

>>>>

\section{Introduction}%Wort ``Path'' durch nicht-definierte Terminologie ersetzen?
Annotating a document is usually solved by adding footnotes or figures in an appropriate position and adding a simple reference in the text, leaving the reader to find the referenced content by themselves. Sometimes however, if a more obvious connection between the text and the referenced content is required, the reference is visibly connected to the text by drawing a straight line or a more complex path between them.

@@

\section{Terminology and Fundamentals} %Evtl in Introduction eingliedern?
While Boundary Labeling (or an equivalent concept) can be applied to a space with different geometry or more dimensions, this paper will only concern itself with two-dimensonal, euclidean space.
To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)%Add references to all terminology pictures!

%Add figure for graph terminology.

@@

\begin{itemize}
 \item \textbf{Graph:} A Set of Vertices and Edges (see below), often used to represent routing problems. Can be further classified depending on its properties.
 \item \textbf{Edge:} A connection between two vertices, usually represented as a line. Can be directional, limiting its usage, or have a cost (``Weight'') associated with its use.
 \item \textbf{Vertex:} A featureless object that can be connected to any number other vertices via edges. Usually depicted as a point or small shape. Also referred to as \textbf{``Node''}.%Remove second sentence and re-define node?
 \item \textbf{Path:} An ordered series of either edges that share one vertex each with the edges listed before and after them, or an ordered series of vertices that have an edge connecting each vertex to the subsequent one. Also known as \textbf{``Graph Walk''}.

 \item \textbf{Polyline:} A connected series of straight lines.%Ist mehr Information notwendig?

 \item \textbf{Label:} Additional information, usually represented as a box containing the additional information. Will also be referred to as \textbf{``Annotation''} in this paper.
 \item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.
 \item \textbf{Leader:} The line or polyline connecting the site to the label - depending on its shape, it can be described further using several subgroups which can be freely combined to form new leader types:
  \begin{itemize}
   \item \textit{S-Leader:} This Leader connects site and label in a straight line.
   \item \textit{O-Leader:} This leader runs orthogonally to the border between text and label area.
   \item \textit{P-Leader:} This Leader runs parallel to the border between text and label area. It must be combined with other leader styles, as it won't reach the label area otherwise.
  \end{itemize}
    For example, the leader from Fig.~\ref{fig:term} would be classified as an opo-Leader.
 \item \textbf{Port:} The location where the leader connects to the label. It may be restricted to pre-determined locations, like only at the corners.
  \item \textbf{Monotonous/Monotonicity:} Describes a steady progression towards a direction or goal. Will be used in this paper to indicate that there's no loss of progress made from the site to the label, effectively restricting the direction a leader can go at each bend.
  \item \textbf{Depth-first search:} A searching algorithm that continues along a set of rules as far as possible until either the goal is found, or backtracking is required. A common example is navigating a maze by turning the same way whenever possible.
\end{itemize}
%Mehr Graphenterminologie? (zb: Path) +Polylinien Vertex/Node für Knoten in Polylinien bzw. Graphen -->Eindeutig!

@@

%Port-Location erwähnen!
\subsection{Algorithm}%''Underlying Algorithm''?
To easily create leaders that exclusively use the space between words and lines, we decided to use a graph similar to the one Loose \cite{Loose2015} used, placing vertices above and below each word's beginning and end, and connecting them with their horizontally and vertically adjacent neighbours.%Umformulieren?
For the sites, we inserted an additional vertex above the center of the word, which will serve as the leader's starting point. We also decided to give the graph's edges a capacity of 1, which means that each leader will always use edges which haven't already be taken by other leaders. While this causes some sites to become un-routable due to another leader passing right through their starting vertex, it makes visualizing the leaders much easier.

Our first approach was going through the labels in the order they appear in the text and placing each as far up as possible, to maximize the remaining space other labels could use. We also opted to use fixed ports located in the top left corner of the annotation, as this allowed us to unambiguously place any annotation by knowing either their port's location, wich is also the leader's ending point.
However, since putting the labels right next to the text would result in the labels' positions being restricted by the text's line spacing, we then decided to leave a buffer zone in between the two areas, which allows the leaders to adjust between the position forced by the line spacing and the label's final position. As this allowed us to place the labels more freely, we adjusted the label's positions after all labels were placed to reduce the distance that needed to be covered in the buffer area.

\subsection{Implementation}
The program was written in Java, using only JGraphT\cite{JGraphT} as additional library. Since we only want to create leaders that don't intersect with the text, the graph was created alongside the placement of the words on the canvas, as it was easy to extract measurements at this point. Due to some problems with Java's various methods of calculating linebreaks, this process was done completely by hand.

As the Graph's vertices represent fixed locations on the canvas, they each have coordinates associated with them, with the vertices representing the sites containing extra information, such as references to the corresponding annotation and to the leader connecting the two (if existent). This additional information is used for both routing and drawing the leaders: The routing was done via a depth-first search algorithm that prioritized vertices located further above if possible, and vertices to the right otherwise. After the algorithm terminates, it returns  a vertex-based graph walk describing the route from the site to the border of the text area alongside information about where to draw the OPO-segment in the buffer zone if successful, or a walk containing only the starting vertex, if no valid path was found. Using this information alongside the positional data stored in the vertices, the polyline representing the leader can be easily drawn.
%TODO: Schönheitsfeatures, etc.
%TODO: Code genauer durchlesen und erweitern!

% For demonstration purposes, we also included another algorithm that uses S-Leaders for this section.
